---
title: "Writing data to googlesheets from R"
output: github_document
---

```{r setup, include=FALSE, message = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Who you are

You use R for data manipulation and often share your data with others using googlesheets. Currently you write the data to an excel/csv file first and then open that file on google sheets which you share with others. 

#### How this document will help you

This document will help you bring down the two-step process to one-step, by writing the data directly to googlesheets. 

## Step 1: Setting up googlesheets package 


```{r gsheetSetup1, eval = FALSE, tidy = TRUE, message = FALSE}

# install and load this package
install.packages("googlesheets")
library(googlesheets)

# Other packages used in this document
library(dplyr)  #For piping
```
## Step 2: Giving access to your google account

```{r gsheetSetup2, eval = FALSE, tidy = TRUE, message = FALSE}
# trigger googlesheets from R console
gs_ls() 

# This command is used to list all the googlesheets you have access to. I am just using this to trigger your gmail account to ask you to grant permission. You can use any command from the googlesheets package here instead!

```

## Step 3: Writing data directly to googlesheets


```{r gsheetWrite1, eval = FALSE, tidy = TRUE, message = FALSE}

# Set your working directory
setwd("~/Google Drive/Tutorials")

# To write a new google sheet
gsheet <- gs_new(title = "spreadsheetTitle", ws_title = "firstSheet", input = head(mtcars),
                    trim = TRUE, verbose = FALSE)

```

Let me break down the `gs_new` command for you:

    title = "spreadsheetTitle" : the name you want to give for your workbook (the googlesheet)
    ws_title = "firstSheet"    : the name of the sheet inside the workbook
    input = head(mtcars)       : this is where you supply the data.frame you want to write


You should now be able to see a googlesheet named "spreadsheetTitle" in your working directory. 


&nbsp;
&nbsp;
&nbsp;




#### To write a new sheet in the same workbook

It is possible to write multiple sheets in a single workbook. You can accomplish that from R using `gs_ws_new`
```{r gsheetWrite2, eval = FALSE, tidy = TRUE, message = FALSE}

gsheet <- gsheet %>%   
  gs_ws_new(ws_title = "secondSheet", input = tail(mtcars),
            trim = TRUE, verbose = FALSE)

```
Let me break down the `gs_ws_new` command for you: 

    gsheet %>%                  : we are using the piping (chaining) operator from dplyr here
    ws_title = "secondSheet"    : the name of the second sheet inside the existing workbook
    input = tail(mtcars)        : the data.frame you want to write


